Joint modelling

### 2020-09-16 manuscript/competing_risks.Rmd
Started writing manuscript
First usable set of hypothesis results

```{r}
hypothesis_res_all <- do.call(rbind, hypothesis_res_list)
write_csv(hypothesis_res_all, path = paste0(cache_dir, "/competing_risks_res.csv"))
```

@@

>>>>

# Peak IL-6, peak D-dimer

```{r}
data_surv_IL_6 <- data_surv %>%
  group_by(patient_id) %>%
  filter(any(!is.na(IL_6))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "IL_6", new_column_name = "peak_IL_6", initial_value = 1) %>%
  mutate(log_peak_IL_6 = log(peak_IL_6))

data_surv_IL_6 %>% group_by(patient_id) %>% summarise(m_b = max(breathing)) %>% group_by(m_b) %>% summarise(n())

f_competing_IL_6 <- Surv(daym1, day,status) ~ log_peak_IL_6

fit_competing_IL_6 <- coxph(f_competing_IL_6, data = data_surv_IL_6, id = patient_id)
fit_competing_IL_6

# hypothesis_res_list[["competing_hcq_az_comorbidities_sum"]] <-
#   rbind(
#   frequentist_hypothesis_res_from_coxph(hypotheses$hcq_reduces_death, fit_competing_hcq_az_comorbidities_sum,
#                                         coefficient_name = "took_hcqTRUE", transition = "1:3", adjusted = "age, sex, az, comorbidities (sum)"),
#   frequentist_hypothesis_res_from_coxph(hypotheses$hcq_increases_discharged, fit_competing_hcq_az_comorbidities_sum,
#                                         coefficient_name = "took_hcqTRUE", transition = "1:2", adjusted = "age, sex, az, comorbidities (sum)")
#   )
# hypothesis_res_list[["competing_hcq_az_comorbidities_sum"]]
```
```{r}
data_surv_d_dimer <- data_surv %>%
  group_by(patient_id) %>%
  filter(any(!is.na(d_dimer))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "d_dimer", new_column_name = "peak_d_dimer", initial_value = 0) %>%
  mutate(log_peak_d_dimer = log(peak_d_dimer + 1) - 4)

data_surv_d_dimer %>% group_by(patient_id) %>% summarise(m_b = max(breathing)) %>% group_by(m_b) %>% summarise(n())

f_competing_d_dimer <- Surv(daym1, day,status) ~ log_peak_d_dimer

fit_competing_d_dimer <- coxph(f_competing_d_dimer, data = data_surv_d_dimer, id = patient_id)
fit_competing_d_dimer

# hypothesis_res_list[["competing_hcq_az_comorbidities_sum"]] <-
#   rbind(
#   frequentist_hypothesis_res_from_coxph(hypotheses$hcq_reduces_death, fit_competing_hcq_az_comorbidities_sum,
#                                         coefficient_name = "took_hcqTRUE", transition = "1:3", adjusted = "age, sex, az, comorbidities (sum)"),
#   frequentist_hypothesis_res_from_coxph(hypotheses$hcq_increases_discharged, fit_competing_hcq_az_comorbidities_sum,
#                                         coefficient_name = "took_hcqTRUE", transition = "1:2", adjusted = "age, sex, az, comorbidities (sum)")
#   )
# hypothesis_res_list[["competing_hcq_az_comorbidities_sum"]]
```

@@

```{r}
hypothesis_res_all <- do.call(rbind, hypothesis_res_list)
write_csv(hypothesis_res_all, path = here::here("manuscript", "competing_risks_res.csv"))
```

### 2020-09-16 manuscript/hmm.Rmd
First usable set of hypothesis results

@@

counterfactual_max_days <- 90
counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
  inner_join(data$patient_data, by = c("patient_id")) %>% 
  left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
  ungroup()

counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)

hcq_m1_counterfactual_took_hcq <- hcq_m1
hcq_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq

hcq_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_took_hcq)

hcq_m1_counterfactual_no_hcq <- hcq_m1
hcq_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq

hcq_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_no_hcq)

@@

```{r}
hcq_m1_epred_rect <- posterior_epred_rect(hcq_m1_fit)
hcq_m1_predicted_rect <- posterior_epred_to_predicted(hcq_m1_fit, hcq_m1_epred_rect)
hcq_m1_predicted <- posterior_rect_to_long(hcq_m1_fit, hcq_m1_predicted_rect)
```

```{r}
hcq_m1_predicted_df <- posterior_long_to_df(hcq_m1_fit$data, hcq_m1_predicted)

@@

```{r}
pp_check_last_state(hcq_m1_fit, hcq_m1_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect)
pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect, states_from = 2:6)
```

@@

@@

summary(hcq_m3_fit$brmsfit)
```
```{r}
hcq_m3_epred_rect <- posterior_epred_rect(hcq_m3_fit)
hcq_m3_predicted_rect <- posterior_epred_to_predicted(hcq_m3_fit, hcq_m3_epred_rect)
hcq_m3_predicted <- posterior_rect_to_long(hcq_m3_fit, hcq_m3_predicted_rect)
```

@@

@@

@@

@@

@@

>>>>

do_pp_checks <- FALSE
hmm_hypothesis_res_list <- list()

@@

# counterfactual_max_days <- 90
# counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
#   inner_join(data$patient_data, by = c("patient_id")) %>% 
#   left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
#   mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
#   mutate(.time = .time +1) %>% 
#   filter(.time >= 1) %>%
#   group_by(.serie) %>%
#   mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
#   ungroup()
# 
# 
# counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# hcq_m1_counterfactual_took_hcq <- hcq_m1
# hcq_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# hcq_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_took_hcq)
# 
# hcq_m1_counterfactual_no_hcq <- hcq_m1
# hcq_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# hcq_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_no_hcq)

@@

```{r}
draws <- tidybayes::tidy_draws(hcq_m1_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m1_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "supportive"
  )

hmm_hypothesis_res_list[["hcq_m1_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "supportive"
  )
```

```{r}
if(do_pp_checks) {
  hcq_m1_epred_rect <- posterior_epred_rect(hcq_m1_fit)
  hcq_m1_predicted_rect <- posterior_epred_to_predicted(hcq_m1_fit, hcq_m1_epred_rect)
  hcq_m1_predicted <- posterior_rect_to_long(hcq_m1_fit, hcq_m1_predicted_rect)
}
```

```{r}
if(do_pp_checks) {
  hcq_m1_predicted_df <- posterior_long_to_df(hcq_m1_fit$data, hcq_m1_predicted)

@@

  }
}
```

```{r}
if(do_pp_checks) {
  pp_check_last_state(hcq_m1_fit, hcq_m1_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
  pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect)
  pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect, states_from = 2:6)
}
```

@@

```{r}
draws <- tidybayes::tidy_draws(hcq_m2_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m2_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive"
  )

hmm_hypothesis_res_list[["hcq_m2_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive"
  )
```

@@

summary(hcq_m3_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(hcq_m3_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m3_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital"
  )

hmm_hypothesis_res_list[["hcq_m3_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital"
  )
```

```{r}
hcq_m3_epred_rect <- posterior_epred_rect(hcq_m3_fit)
hcq_m3_predicted_rect <- posterior_epred_to_predicted(hcq_m3_fit, hcq_m3_epred_rect)
hcq_m3_predicted <- posterior_rect_to_long(hcq_m3_fit, hcq_m3_predicted_rect)
```

@@

```{r}
draws <- tidybayes::tidy_draws(hcq_m4_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m4_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital, az"
  )

hmm_hypothesis_res_list[["hcq_m4_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital, az"
  )

hmm_hypothesis_res_list[["hcq_m4_death_az"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_azTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital, hcq"
  )

hmm_hypothesis_res_list[["hcq_m4_discharged_az"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_azTRUE + draws$`b_rate_groupimprove_one:took_azTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital, hcq"
  )
```

@@

```{r}
draws <- tidybayes::tidy_draws(hcq_m5_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m5_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital, comorbidities_sum"
  )

hmm_hypothesis_res_list[["hcq_m5_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital, comorbidities_sum"
  )
```

@@

```{r}
draws <- tidybayes::tidy_draws(hcq_m6_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m6_d_dimer_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_log_peak_d_dimer,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$d_dimer_increases_death,
    adjusted = "age, sex, hospital"
  )

@@

```{r}
draws <- tidybayes::tidy_draws(hcq_m7_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m7_IL_6_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_log_peak_IL_6,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$IL_6_increases_death,
    adjusted = "age, sex, hospital"
  )

@@

# Environment

```{r}
hmm_hypothesis_res_all <- do.call(rbind, hmm_hypothesis_res_list)
write_csv(hmm_hypothesis_res_all, path = here::here("manuscript", "hmm_res.csv"))
```

```{r}
sessionInfo()
```

### 2020-09-16 manuscript/jm.Rmd
First usable set of hypothesis results

```{r setup}
devtools::load_all()
library(tidyverse)
library(survival)
library(rstanarm)

@@

@@

data_d_dimer %>% filter(d_dimer == 0)
# data_d_dimer %>% group_by(sex, ever_hcq) %>% summarise(pts = length(unique(patient_id)), n_dimer = sum(!is.na(d_dimer)))

fit_d_dimer_age_sex_hcq <- stan_jm_with_cache(formulaLong = log_d_dimer ~ day + (1  | patient_id),
              dataLong = data_d_dimer %>% filter(!is.na(d_dimer)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex + took_hcq,
              dataEvent = data_d_dimer,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_d_dimer_age_sex_hcq.rds"))

@@

#length(unique(data_d_dimer$patient_id))
to_pred <- data_d_dimer %>% group_by(patient_id) %>% summarise(n_d_dimer = sum(!is.na(d_dimer))) %>%
  filter(n_d_dimer > 1) %>% pull(patient_id)
#data_d_dimer %>% filter(patient_id == unique(data_d_dimer$patient_id)[31])

```
```{r}
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[1:11])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[12:14])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[15:18])
plot(p1, plot_observed = TRUE)

@@

fit_IL_6_age_sex_hcq <- stan_jm_with_cache(formulaLong = log_IL_6 ~ day + (1 + day | patient_id),
              dataLong = data_IL_6 %>% filter(!is.na(IL_6)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex + took_hcq,
              dataEvent = data_IL_6,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_IL_6_age_sex_hcq.rds"))

fit_IL_6_age_sex_hcq

```
```{r}
for(id in unique(data_IL_6$patient_id)) {
  p1 <- posterior_traj(fit_IL_6_age_sex_hcq, ids = id)
  print(plot(p1, plot_observed = TRUE)                     )
}

@@

>>>>

```{r setup}
devtools::load_all()
library(tidyverse)
library(survival)
library(rstanarm)
library(tidybayes)

@@

jm_hypothesis_res_list <- list()

@@

#data_d_dimer %>% filter(d_dimer == 0)
# data_d_dimer %>% group_by(sex, ever_hcq) %>% summarise(pts = length(unique(patient_id)), n_dimer = sum(!is.na(d_dimer)))
fit_d_dimer_age_sex <- stan_jm_with_cache(formulaLong = log_d_dimer ~ day + (1  | patient_id),
              dataLong = data_d_dimer %>% filter(!is.na(d_dimer)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex,
              dataEvent = data_d_dimer,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_d_dimer_age_sex.rds"))

fit_d_dimer_age_sex

#length(unique(data_d_dimer$patient_id))
# to_pred <- data_d_dimer %>% group_by(patient_id) %>% summarise(n_d_dimer = sum(!is.na(d_dimer))) %>%
#   filter(n_d_dimer > 1) %>% pull(patient_id)
#data_d_dimer %>% filter(patient_id == unique(data_d_dimer$patient_id)[31])

```{r}
jm_hypothesis_res_list[["d_dimer_age_sex"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$d_dimer_increases_death,
    fit_d_dimer_age_sex,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex")

jm_hypothesis_res_list[["d_dimer_age_sex"]]
```

```{r}
fit_d_dimer_age_sex_hcq <- stan_jm_with_cache(formulaLong = log_d_dimer ~ day + (1  | patient_id),
              dataLong = data_d_dimer %>% filter(!is.na(d_dimer)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex + took_hcq,
              dataEvent = data_d_dimer,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_d_dimer_age_sex_hcq.rds"))

@@

#length(unique(data_d_dimer$patient_id))
# to_pred <- data_d_dimer %>% group_by(patient_id) %>% summarise(n_d_dimer = sum(!is.na(d_dimer))) %>%
#   filter(n_d_dimer > 1) %>% pull(patient_id)
#data_d_dimer %>% filter(patient_id == unique(data_d_dimer$patient_id)[31])

```{r}
jm_hypothesis_res_list[["d_dimer_age_sex_hcq"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$d_dimer_increases_death,
    fit_d_dimer_age_sex_hcq,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex, hcq")

jm_hypothesis_res_list[["d_dimer_age_sex_hcq"]]
```

```{r}
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[1:11])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[12:14])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[15:18])
plot(p1, plot_observed = TRUE)

@@

fit_IL_6_age_sex <- stan_jm_with_cache(formulaLong = log_IL_6 ~ day + (1 + day | patient_id),
              dataLong = data_IL_6 %>% filter(!is.na(IL_6)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex,
              dataEvent = data_IL_6,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_IL_6_age_sex.rds"))

fit_IL_6_age_sex

```
```{r}
jm_hypothesis_res_list[["IL_6_age_sex"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$IL_6_increases_death,
    fit_IL_6_age_sex,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex")

jm_hypothesis_res_list[["IL_6_age_sex"]]

@@

```{r}
fit_IL_6_age_sex_hcq <- stan_jm_with_cache(formulaLong = log_IL_6 ~ day + (1 + day | patient_id),
              dataLong = data_IL_6 %>% filter(!is.na(IL_6)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex + took_hcq,
              dataEvent = data_IL_6,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_IL_6_age_sex_hcq.rds"))

fit_IL_6_age_sex_hcq
```

```{r}
jm_hypothesis_res_list[["IL_6_age_sex_hcq"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$IL_6_increases_death,
    fit_IL_6_age_sex_hcq,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex, hcq")

jm_hypothesis_res_list[["IL_6_age_sex_hcq"]]

```{r}
for(id in unique(data_IL_6$patient_id)) {
  p1 <- posterior_traj(fit_IL_6_age_sex_hcq, ids = id)
  print(plot(p1, plot_observed = TRUE)                     )
}

```{r}
jm_hypothesis_res_all <- do.call(rbind, jm_hypothesis_res_list)
write_csv(jm_hypothesis_res_all, path = here::here("manuscript", "jm_res.csv"))
```
